---
title: "List 1"
author: "Dawid Dieu"
date: '2022-03-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Generate data
```{r}
set.seed(2022)
generate_data <- function() {
  X = matrix(
    data=rnorm(n=1000 * 950, mean=0, sd = 1 / sqrt(1000)), 
    nrow=1000, 
    ncol=950,
  )
  
  Beta = rep(x=0, times=950)
  Beta[0: 5] = 3
  
  eps = rnorm(n=1000, mean=0, sd=1)
  
  Y = X %*% Beta + eps
  
  return (list(X, Y))
}

```

### Perform the following analysis using reduced models
```{r}
set.seed(2022)
print_fdr_info <- function(tp, fp, prefix) {
  cat(prefix, 'True Positives: ', tp , '\n')
  cat(prefix, 'False Positives: ', fp , '\n')
  fdr = fp / (fp + tp + 10^-9) * 100
  cat(prefix, 'False discovery rate (FDR): ', fdr, '%\n')
}


run_test <- function(show_debug=TRUE) {
  n_features = c(10, 100, 500, 950)
  new_data = generate_data()
  X = new_data[[1]]
  Y = new_data[[2]]
  
  for (n in n_features) {
    if(show_debug) cat('\n📊Test for', n, 'columns\n')
    if(show_debug) cat('Shape of X:', dim(X[, 1:n]), '\n')
    
    model = lm(formula=Y ~ X[, 1:n])
    Beta_hat = model $ coefficients  # least squares estimator of Beta from lm
    p_values = summary(model) $ coefficients[, 4]
    
    # add vector of 1s which are coefficients of the intercept
    new_X = cbind(rep(x=1, times=1000), X[, 1:n])
    
    # least squares estimator of Beta from closed form solution
    myBeta = solve(a = t(new_X) %*% new_X) %*% t(new_X) %*% Y
    
    diff = sum((Beta_hat - myBeta) ^ 2)
    cat('Beta (my manual calculation) vs Beta from lm SSE difference: ', diff, '\n')
    
    #  a) tests for significance of individual regression coefficients
    alpha = 0.1
    discoveries = p_values < alpha
    cat('🔸a) Number of significant Beta_hat: ', sum(discoveries), '\n')
  
    # b) average standard deviation of the estimators of individual regression coefficients
    Y_hat = new_X %*% Beta_hat
    SSE = t((Y - Y_hat)) %*% (Y - Y_hat)
    dfe = 1000 -  n # degrees of freedom
    MSE = (SSE / dfe)[1]
    cov_matrix = MSE * solve(t(new_X) %*% new_X)
    sigmas = sqrt(diag(cov_matrix))
    cat('🔹b) Average standard deviation (my manual calculation): ', mean(sigmas), '\n')
    
    sigmas_from_summary = summary(model) $ coefficients[, 2]
    cat('🔹b) Average standard deviation (from model summary): ', mean(sigmas), '\n')
    
    # average length of the 90% confidence intervals (theoretical estimate)
    t = qt(p = 1 - alpha / 2, df = 1000 - n)
    interval = t * sigmas * 2
    cat('🔹b) Average length of the 90% confidence intervals: ', mean(interval), '\n')
    
    # True and False discoveries
    true_positives = sum(discoveries[2:6])
    # we add discoveries[1] because we don't use intercept in our model
    false_positives = as.integer(discoveries[1])  + sum(discoveries[7:length(discoveries)])
    cat('🔸c) i) Without adjusting for multiple testing\n')
    print_fdr_info(tp=true_positives, fp=false_positives, prefix='🔸c) i)')
    
    # sort p values
    sorted_indices = order(p_values)
    
    # Bonferroni correction; reject i-th coefficient if p < alpha / n
    cat('🔸c) ii) Bonferroni correction\n')
    bonferroni_rejected_indices = sorted_indices[p_values[sorted_indices] < (alpha / n)]
    true_positives = sum(bonferroni_rejected_indices >= 2 & bonferroni_rejected_indices <= 6)
    false_positives = sum(bonferroni_rejected_indices == 1 & bonferroni_rejected_indices >= 7)
    print_fdr_info(tp=true_positives, fp=false_positives, prefix='🔸c) ii) [Bonferroni]')
  
    # Benjamini-Hochberg correction; reject i-th coefficient if p_i < i * alpha / n
    cat('🔸c) iii) Benjamini-Hochberg correction\n')
    bh_rejected_indices = sorted_indices[
      p_values[sorted_indices] < (alpha / n) * seq(from=1, to=n + 1, by=1)
    ]
    true_positives = sum(bh_rejected_indices >= 2 & bh_rejected_indices <= 6)
    false_positives = sum(bh_rejected_indices == 1 & bh_rejected_indices >= 7)
    print_fdr_info(tp=true_positives, fp=false_positives, prefix='🔸c) iii) [Benjamini-Hochberg]')
  }
}

run_test()
```
## Problem 1

### a)
#### Number of significant coefficients for each model with $\alpha=0.1$

| number of columns | number of significant coefficients |
|:-----------------:|:-------------------------------:|
|         10        |                4                |
|        100        |                11               |
|        500        |                45               |
|        950        |               105               |

### b)
#### Average standard deviation of the estimators of individual regression coefficients and the average length of the respective 90% confidence intervals

| number of columns | average standard deviation | average length of 90% confidence intervals |
|:-----------------:|:--------------------------:|:------------------------------------------:|
|         10        |          0.9087013         |                  2.992161                  |
|        100        |          1.038993          |                  3.421506                  |
|        500        |           1.39654          |                  4.602734                  |
|        950        |          4.366425          |                  14.63543                  |

### c)
#### False Discoveries Correction
##### i) Without adjusting for multiple testing

| number of columns | True Positives | False Positives |   FDR  |
|:-----------------:|:--------------:|:---------------:|:------:|
|         10        |        4       |        0        |   0%   |
|        100        |        5       |        6        | 54.54% |
|        500        |        2       |        43       | 95.55% |
|        950        |        1       |       104       | 99.04% |


##### ii) using Bonferroni correction

| number of columns | True Positives | False Positives | FDR |
|:-----------------:|:--------------:|:---------------:|:---:|
|         10        |        2       |        0        |  0% |
|        100        |        1       |        0        |  0% |
|        500        |        0       |        0        |  0% |
|        950        |        0       |        0        |  0% |


##### ii) using Benjamini-Hochberg correction

| number of columns | True Positives | False Positives | FDR |
|:-----------------:|:--------------:|:---------------:|:---:|
|         10        |        2       |        0        |  0% |
|        100        |        1       |        0        |  0% |
|        500        |        0       |        0        |  0% |
|        950        |        0       |        0        |  0% |



## Problem 2



